### 2020-11-30 paper/2020-10-revision-background.tex
Introduce female education and associated effects

Holding son preference constant, lower desired fertility increases the use of sex 
selection, suggesting that the first order effect of higher female wages is to increase 
sex selection because of the fertility effect discussed above \citep{Portner2015b,Jayachandran2017}.  
Furthermore, for a given desired number of children, the higher the parity the more 
likely the use of sex selection if the desired number of sons has not been reached.

% Factors/predictions

>>>>

Holding son preference constant, lower desired fertility increases the use of sex 
selection, suggesting that the first order effect of higher female wages is to increase 
sex selection because of the fertility effect discussed above \citep{Portner2015b,Jayachandran2017}.  
Furthermore, the higher the parity the more likely the use of sex selection if the 
desired number of sons has not been reached.
[cost of sex selection?]

% Factors/predictions
An important issue is how to operationalize the potential explanations of how 
birth spacing has changed over time.
Although the explanations are based on changes in wages, mortality risk, degree of son 
preference, and access to and cost of sex selection, none of these are available for the 
period when the decisions of interest were made.

Two characteristics that are available in data and that have been shown to be associated 
with fertility, child outcomes, and the use of sex selection are female education and 
area of residence.
[citations?]
The problem is that both have multiple pathways through which they can affect fertility,
birth spacing, and sex selection and that these do not necessarily line up neatly with the
theories above.

Families in urban areas are likely to have fewer children and higher rates of sex 
selection and, therefore, increasingly longer birth intervals over time compared to
rural families.
With the costs of children higher in urban than in rural areas both because of the
lower opportunities for productive work of children and the higher cost of consumption
related to children, fertility is expected to be lower in urban than in rural areas.
Furthermore, access to health care and generally a healthier environment because of,
for example, cleaner water means that urban parents can have fewer children and invest
more in each than in rural areas.%
\footnote{
However, holding parental education constant, there is evidence that children in
Indian slums do worse than they would have in rural areas \citep{Portner2018a}.
}
However, the better access to health care also means that there is easier access to
sex selection and the generally higher income of urban households makes it more
affordable.
[differences in birth intervals without sex selection?]

Wages are generally increasing in education suggesting that with higher female education
we should see shorter birth spacing and lower fertility, but also higher use of sex 
selection.
However, female education undoubtedly represents more than the value of their time.
Two examples are child health and the production of offspring human capital.

First, children of more educated women tend to have better health outcomes and
lower mortality than less educated women, which may come from higher income, but could
also be a result of better health knowledge 
\citep{Rosenzweig1982a,Kovsted2002,Whitworth2002,Maitra2008}.
Lower expected mortality means that fewer births are required to reach the desired
number of surviving children.
Similarly, although short birth spacing generally leads to worse health outcomes, 
the more educated the mother, the more muted this relationship is, which, at least partly,
removes the health outcome incentive for longer birth spacing
\citep{Whitworth2002,Conde-Agudelo2012,Molitoris2019}.

Second, more educated women are more efficient at producing child human capital 
\citep{Behrman1999}.
Hence, families that value sons' education more will both tend to demand better-educated
women as mothers and to have fewer children because both allow for higher quality boys.
With the increasing return to male education that India has experienced, this imply
that there will be an increased demand for better-educated women combined with lower 
fertility, and an increased use of sex selection when available.
In this case, the associations between higher female education, lower fertility, and more
sex selection are not a causal effect driven by higher wages but instead the result of the 
increased return to male education.
